Advancing Renewable Gas Storage using Flat-histogram Methods

### 2020-06-08 dissertation/introduction.tex
made all committee changes; now working on putting appendix into chapter.

\section{Canonical Ensemble}
The Canonical ensemble is a representation for a system $\Gamma$ that is
determined by a fixed number of particles $N$ (and fixed components $\sum N_i$
for a multi-component system), a fixed volume $V$, and a fixed temperature $T$.
All possible states for the system $\Gamma\left(N,V,T\right)$ have the same
number of particles, energy, and temperature due to the system being in thermal
equilibrium with a heat bath. The canonical ensemble is for this reason also
known as the $NVT$ ensemble.

@@

\subsection{Square-Well Fluid}
The square-well fluid is a system of particles whose interactions are governed by a square-well
potential~\cite{singh2003surface, barker2004perturbationSW}.  The square-well potential is an ideal test-bed
for Monte Carlo methods as it is a simple model for a liquid, which includes both attractive and repulsive
interactions~\cite{barker1967-SW-perturbation, vega1992phase}. The potential $U(\textbf{r})$ for such a system
is given by
\begin{equation}
 U(\textbf{r})=\begin{cases} \infty &
 \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
 \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
 \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
\end{equation}
where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the reduced range of the potential well, and $\epsilon$ is its depth.

\subsection{2D Ising Model}
The 2D Ising spin-lattice system is widely used as a test-bed when benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The \nth{2} order phase
transition behavior and the ability to directly calculate the exact solution for finite
lattices~\cite{beale1996exact} make the system sufficiently interesting for such theoretical comparisons.
The Hamiltonian for the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new} cab be written as the following:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i \sigma_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the absence of a magnetic field ($h =
0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over nearest neighbor spin sites.

\subsection{Lennard-Jones Clusters}
Lennard-Jones cluster systems are widely used as a challenging test-bed when comparing Monte-Carlo
methods~\cite{doye1998thermodynamics, poulain2006performances, martiniani2014superposition}. The potential
energy for interacting atomic particles can be written as follows:
\begin{align}
U = 4\epsilon \left[\left( \sigma/r \right)^{12} - \left( \sigma/r \right)^{6}\right]
\end{align}
A number of challenges exist when trying to accurately simulate the density of states of a cluster system.
Narrow potential energy funnels~\cite{mandelshtam2006multiple, wales1997global} and low temperature solid-solid
transitions~\cite{neirotti2000phase, calvo2000phase, calvo2000entropic} make predicting thermodynamic properties
such as heat capacity an extremely difficult process. In addition, since flat-histogram methods explore all
energies and all temperatures it is possible that for high temperatures a large cluster volume is possible
(which could lead to boil-off of the particles). A `constraining' sphere with radius $R_c$ is used to prevent
this but it is usually empirically determined, further complicating the simulation process.

>>>>

\section{Canonical Ensemble}
The Canonical ensemble is a representation for a system $\Gamma$ that is
determined by a fixed number of particles $N$ (and fixed components $\sum N_i$
for a multi-component system), a fixed volume $V$, and a fixed temperature $T$.
All possible states for the system $\Gamma\left(N,V,T\right)$ have the same
number of particles, volume, and temperature due to the system being in thermal
equilibrium with a heat bath. The canonical ensemble is for this reason also
known as the $NVT$ ensemble.

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/well-potential.pdf}
    \caption{The square-well and the Lennard-Jones potentials are graphically shown vs. the radial distance from the particle.
    }
    \label{fig:well-potential}
  \end{figure}

\subsection{Square-Well Fluid}
The square-well fluid is a system of particles whose interactions are governed by a square-well
potential~\cite{singh2003surface, barker2004perturbationSW} as shown in Fig.~\ref{fig:well-potential}.  The square-well potential is an ideal test-bed
for Monte Carlo methods as it is a simple model for a liquid, which includes both attractive and repulsive
interactions~\cite{barker1967-SW-perturbation, vega1992phase}. The potential $U(\textbf{r})$ for such a system
is given by
\begin{equation}
 U(\textbf{r})=\begin{cases} \infty &
 \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
 \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
 \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
\end{equation}
where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the reduced range of the potential well, and $\epsilon$ is its depth.

\subsection{2D Ising Model}
The 2D Ising spin-lattice system is widely used as a test-bed when benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The \nth{2} order phase
transition behavior and the ability to directly calculate the exact solution for finite
lattices~\cite{beale1996exact} make the system sufficiently interesting for such theoretical comparisons.
The Hamiltonian for the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new} cab be written as the following:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i \sigma_i
\end{align}
The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
for up or down spins respectively. In the absence of a magnetic field ($h =
0$), we can write the Hamiltonian as follows~\cite{onsager1944crystal,
kaufman1949crystal}:
\begin{align}
\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
\end{align}
where the sum is over nearest neighbor spin sites.

\subsection{Lennard-Jones Clusters}
Lennard-Jones cluster systems are widely used as a challenging test-bed when comparing Monte-Carlo
methods~\cite{doye1998thermodynamics, poulain2006performances, martiniani2014superposition}. The potential
energy for interacting atomic particles can be written as follows and can visually be seen in Fig.~\cite{fig:well-potential}:
\begin{align}
U = 4\epsilon \left[\left( \sigma/r \right)^{12} - \left( \sigma/r \right)^{6}\right]
\end{align}
A number of challenges exist when trying to accurately simulate the density of states of a cluster system.
Narrow potential energy funnels~\cite{mandelshtam2006multiple, wales1997global} and low temperature solid-solid
transitions~\cite{neirotti2000phase, calvo2000phase, calvo2000entropic} make predicting thermodynamic properties
such as heat capacity an extremely difficult process. In addition, since flat-histogram methods explore all
energies and all temperatures it is possible that for high temperatures a large cluster volume is possible
(which could lead to boil-off of the particles). A `constraining' sphere with radius $R_c$ is used to prevent
this but it is usually empirically determined, further complicating the simulation process.
